---
title: "Outcomes after resection for melanoma"
author: "The Melanoma Collaborative"
output:
  html_notebook: default
  pdf_document: default
---

## Introduction
Melanoma is a skin cancer with poor prognosis. In this study we re-analyse a classical dataset of outcomes after melanoma surgery and compare these to contemporary studies. The aim is to characterise the effect of ulceration on outcome, while correcting for other prognostic markers. 

<!-- Load libraries (Ctrl-Shift-C to do this comment)-->

```{r include=FALSE}
# I don't want these to show, so I will run them but have no output. 
library(finalfit)
library(ggplot2)
library(dplyr)
library(forcats)
library(knitr)
library(boot)
load(file = here::here("data_original", "melanoma_recode.rda"))
theme_set(theme_classic())
```

## Methods
Kaplan Meier and Cox Proportional Hazards regression approaches were used to analyse data.We have used RMarkdown to produce a document that includes all the analysis and text required for this report. 

## Results

### Patient characteristics

`r dim(melanoma)[1]` patients were included from between `r min(melanoma$year)` and `r max(melanoma$year)`. Patients were aged `r min(melanoma$age)` to `r max(melanoma$age)` years. A comparison by ulceration status is shown in Table 1. Tumour thickness and sex were associated with tumour ulceration.  


```{r Demographics, echo=FALSE}
dependent = "ulcer"
explanatory = c("age", "sex", "year", "thickness")
melanoma %>% 
  summary_factorlist(dependent, explanatory, p = TRUE) %>% 
  
  # Format table using kable for output. 
  dependent_label(melanoma, dependent) %>% 
  kable(row.names = FALSE, align = c("l", "l", "r", "r", "r"),
        caption = "Demographics.")
```

### Disease-specific survival

Patients were more likely to die of melanoma if the tumour was ulcerated (Figure 1).

```{r Ulceration plots, echo=FALSE, fig.cap="Outcome by tumour ulceration", fig.height=3, fig.width=7}
library(patchwork)

p1 = melanoma %>%
  ggplot() + 
  aes(x = ulcer,
      fill = status) + 
  geom_bar(show.legend = FALSE) + 
  xlab("Ulceration") + 
  ylab("N")

p2 = melanoma %>%
  ggplot() + 
  aes(x = ulcer,
      fill = status) + 
  geom_bar(position = "fill") +
  xlab("Ulceration") + 
  ylab("Proportion")

p1 + p2
```

A Kaplan-Meier analysis showed a clear disease-specific survival difference between patients with ulceration and those without (p<0.0001) (Figure 2). 

```{r Survival plots, echo=FALSE, fig.cap="Survival after melanoma surgery. Disease-specific survival (A) and overall survival (B). ", fig.height=4, fig.width=7}
dependent = "Surv(time/365, status == 'Died melanoma')"
explanatory = "ulcer"
p1 = melanoma %>% 
  surv_plot(dependent, explanatory, xlab = "Time (years)", 
            fontsize = 4, font.tickslab = 10,
            pval = TRUE, pval.size = 4,
            title = "A. Disease-specific survival")

dependent = "Surv(time/365, status == 'Died melanoma' | status == 'Died not melanoma')"
explanatory = "ulcer"
p2 = melanoma %>% 
  surv_plot(dependent, explanatory, xlab = "Time (years)", 
            fontsize = 4, font.tickslab = 10, 
            pval = TRUE, pval.size = 4,
            title = "B. Overall survival")

survminer::arrange_ggsurvplots(list(p1, p2))
```

```{r Regression table - DSS, echo=FALSE}
dependent = "Surv(time, status == 'Died melanoma')"
explanatory = c("ulcer", "age", "sex", "year", "thickness")
melanoma %>% 
  finalfit(dependent, explanatory, add_dependent_label = FALSE) %>% 
  {fit_dss <<- .} %>% 
  rename("Disease-specific survival" = label) %>% 
  rename(" " = levels) %>% 
  kable(row.names = FALSE, align = c("l", "l", "r", "r", "r"),
        caption = "Disease-specific survival. HR hazard ratio, SD standard deviation.")
```

In a univariable analysis, there was a signficant association between ulceration and a worse disease-specific survival (HR `r fit_dss[2,3]`). This persisted in a multivariable analysis (HR `r fit_dss[2,4]`) correcting for age, sex, operation year and tumour thickness (Table 2)(Figure 3).   

```{r HR plot, echo=FALSE, fig.cap="Disease-specific survival after melanoma surgery. HR hazard ratio, CI confidence interval, SD standard deviation", fig.height=3, fig.width=12, warning=FALSE}
dependent = "Surv(time, status == 'Died melanoma')"
explanatory = c("ulcer", "age", "sex", "year", "thickness")
melanoma %>% 
  hr_plot(dependent, explanatory, dependent_label = "Disease-specific survival")
```

### Overall survival

Overall survival was also worse in patients who had an ulcerated melanoma (log-rank test p<0.0001)(Figure 2B). 

```{r echo=FALSE}
dependent = "Surv(time/365, status == 'Died melanoma' | status == 'Died not melanoma')"
explanatory = c("ulcer", "age", "sex", "year", "thickness")
melanoma %>% 
  finalfit(dependent, explanatory, add_dependent_label = FALSE) %>% 
  {fit_os <<- .} %>% 
  rename("Overall survival" = label) %>% 
  rename(" " = levels) %>% 
  kable(row.names = FALSE, align = c("l", "l", "r", "r", "r"),
        caption = "Overall survival. HR hazard ratio, SD standard deviation.")
```

In a univariable analysis, there was also a signficant association between ulceration and overall survival (HR `r fit_os[2,3]`) which was also seen in the multivariable analysis (HR `r fit_os[2,4]`)(Table 3)(Figure 4).   

```{r echo=FALSE, fig.cap="Overall survival after melanoma surgery. HR hazard ratio, CI confidence interval, SD standard deviation", fig.height=3, fig.width=12, warning=FALSE}
dependent = "Surv(time/365, status == 'Died melanoma' | status == 'Died not melanoma')"
explanatory = c("ulcer", "age", "sex", "year", "thickness")
melanoma %>% 
  hr_plot(dependent, explanatory, dependent_label = "Overall survival")
```

## Discussion

In this study we have re-analysed a classical dataset of outcomes after melanoma surgery. We have shown that that tumour ulceration is strong negative prognostic factor and is independent of better appreciated markers such as tumour thickness. 
