---
title: "Outcomes after resection for colorectal cancer"
author: "The Colorectal Cancer Collaborative"
output:
  pdf_document: default
  html_notebook: default
  html_document:
    df_print: paged
---

```{r include=FALSE}
library(tidyverse)
library(finalfit)
library(knitr)
library(boot)
theme_set(theme_classic())
load(file = here::here("processed-data", "colon_s_recode.rda"))
```

## Introduction
Colorectal cancer is the third most common cancer worldwide. In this study, we have re-analysed a classic dataset to determine the influence of nodal disease in overall survival prior to the introduction of modern chemotherapeutic regimes. 

## Methods
Kaplan Meier and Cox Proportional Hazards regression approaches were used to analyse data.We have used RMarkdown to produce a document that includes all the analysis and text required for this report. 

## Results

### Patient characteristics

```{r}
dependent = "node4.factor"
explanatory = c("age", "sex.factor", "extent.factor", "differ.factor")

# Use summary_factorlist()
# Use kable() to format and include a table caption
# Use chunk cog to hide code
```

### Outcome by nodal status

```{r}
colon_s %>%
  ggplot() + 
  aes(x = node4.factor,
      fill = status.factor) + 
  geom_bar() +
  xlab(">4 lymph nodes") + 
  ylab("N")

colon_s %>%
  ggplot() + 
  aes(x = node4.factor,
      fill = status.factor) + 
  geom_bar(position = "fill") +
  xlab(">4 lymph nodes") + 
  ylab("Proportion")

# Combine these into single plot using library(patchwork)
# Add a figure caption in the chunk preamble
# Adjust figure height and width in preamble
# Use chunk cog to hide code
```

Write something about results here.

```{r}
dependent = "Surv(time, status)"
explanatory = "node4.factor"
colon_s %>% 
  surv_plot(dependent, explanatory)

# Add a figure caption in the chunk preamble
# Adjust figure height and width in preamble
# Use chunk cog to hide code
```

### Overall survival

```{r}
dependent = "Surv(time, status)"
explanatory = c("age", "sex.factor", "extent.factor", "differ.factor", "node4.factor")

# Use the function finalfit()
# Use kable() to shape final table
# Use chunk cog to hide code
```

Write something about results here. 

```{r fig.width=10}
# Make an HR plot
# Add a figure caption
# Adjust height and width if necessary
# Use chunk cog to hide code
```
 
## Discussion
Conclude something about results here. 
