---
title: "Colorectal cancer analysis notebook"
output:
  pdf_document: default
  html_notebook: default
---

## Background
Use what you have learned to do some data exploration of a new dataset, `colon_s`. 

These are data from one of the first successful trials of adjuvant chemotherapy for colon cancer. Levamisole is a low-toxicity compound previously used to treat worm infestations in animals; 5-FU is a moderately toxic (as these things go) chemotherapy agent. 

We are particularly interested in the relationship beween lymph node status (`node4.factor`) and survival. This variable is categorical for patients with 4 or more positive lymph nodes found at cancer surgery. 

## Load libraries

```{r}
library(tidyverse)
library(finalfit)
library(knitr)
library(boot)
theme_set(theme_classic())

```

## Get and tidy data
By hitting `F1` on `data(colon_s)` below we can see the codebook for the dataset. 

```{r}
data(colon_s)
```

## Explore data
Look at the relationship between nodal status (`node4.factor`) and some of the other variables. 

### Plotting
Start a chunk. Remember, `Insert` above does this or `Ctrl-Alt-I`. 

Give the plot a title:

#### Outcome by nodal status (count)

```{r}
colon_s %>%
  ggplot() + 
  aes(x = node4.factor,
      fill = status.factor) + 
  geom_bar()
```

#### Outcome by tumour ulceration (proportion)
Using code from `03_analysis_pipeline.Rmd`, complete the following. 

```{r}
```

### Confounding
Is nodal status confounded by other variables that may influence outcome?

```{r}
dependent = "node4.factor"
explanatory = c("age", "sex.factor", "extent.factor", "differ.factor")

# Use summary_factorlist()

```

Write here what the results show. 

## Analysis

### Plotting

```{r}
dependent = "Surv(time, status)"
explanatory = "node4.factor"
colon_s %>% 
  surv_plot(dependent, explanatory)

```
### Model building

#### Univariable with variable of interest

```{r}
# Use the code in the previous chunk with the function finalfit()
```

#### Multivariable including confounders

```{r}
dependent = "Surv(time, status)"
explanatory = c("age", "sex.factor", "extent.factor", "differ.factor", "node4.factor")

# Use the function finalfit()
```

What do the results of the multivariable analysis show? Write these here. Which variables are independently predictive of survival?

### Extracting results to text
Re-create the above chunk but save the table as `fit`. Extract the `HR (multivariable)` for `node4.factor` and put in a sentence. 

### HR plot 

```{r fig.width=10}

```
 
## Preview, PDF and Word document
Use `Run All` in the dropdown top right, then `Knit` your document in the above 3 formats. Make any adjustments to the presentation.  

## Save final dataset
```{r}
save(colon_s, file = here::here("data_original", "colon_s_recode.rda"))
```
